---
title: "Untitled"
author: "Sergio Grueso"
date: "11/10/2019"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<<<<<<< HEAD

# 

# Download al load the libraries and data of the longitudinal and cross sectional studies (cross sectional study will not be used for now).

```{r}
library(tidyverse)
library(dslabs)
library(dbplyr)
library(ggplot2)

set.seed(1, sample.kind = "Rounding")
longitudinal <- read.csv("../MRI/oasis_longitudinal.csv", header = TRUE, stringsAsFactors = FALSE)
set.seed(1, sample.kind = "Rounding")
cross <- read.csv("../MRI/oasis_cross-sectional.csv", header = TRUE, stringsAsFactors = FALSE)
```


# Summary of the longitudinal dataset

```{r}
str(data)

summary(longitudinal)
```


## Data analysis:

# Hand variable is not relevant because it has the same values for all subjects. Hand is droped.


```{r}
longitudinal$Hand <- NULL
```

# M.F variable name is changed to "Sex" to be more descriptive


```{r}
names(longitudinal)[names(longitudinal) == "M.F"] <- "Sex"
```


# Quick easy search for null values

```{r}
sum(is.na(longitudinal))
```

# 19 observations have one missing value from SES variable and 2 observations have missing values from SES and MMSE column. Subjects with 2 missing values are dropped and any missing value is replaced by the median of that variable. Median is used instead of mean because SES is a variable with integer values from 1 to 5 that indicates socioeconomical status and we do not want decimals values.

```{r}
ind <- which(is.na(longitudinal$MMSE))
longitudinal <- longitudinal[-ind, ]

ind <- which(is.na(longitudinal$SES))
longitudinal$SES[ind] <- median(longitudinal$SES[-ind])
```

# Group variable is the variable we will want to predict so demented is changed to 1 and nondemented to 0. This is in order to facilitate predictions, if the algorithm predicts anything greater than 0.5 (>0.5) the prediction will be "1" meaning that pacient has or is in risk of Alzheimer.

```{r}
longitudinal$Group <- ifelse(longitudinal$Group=="Demented", 1, 0)
```


## Data visualization
# Data visualization analysis can be started by observing the Group, Age and Sex distribution

```{r}
longitudinal %>% ggplot(aes(Group, Age, fill= Sex)) +
    geom_boxplot()
```

# Results on the test MMSE plotted against Age and colored depending on the Group.

```{r}
longitudinal %>% ggplot(aes(Age, MMSE, col = Group)) +
    geom_point()
```

#
































































































=======
```{r}
set.seed(1, sample.kind = "Rounding")
data <- read.csv("../MRI/oasis_longitudinal.csv", header = TRUE, stringsAsFactors = FALSE)
set.seed(1, sample.kind = "Rounding")
data_cross <- read.csv("../MRI/oasis_cross-sectional.csv", header = TRUE, stringsAsFactors = FALSE)
```

>>>>>>> c9dc1cb12a8cf4151c5e0d92d7e5e3abaa877b5f
